NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-29 paper/application.tex
update section 6.4

\begin{figure}[htbp]
\centering
%\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples). The user can drill down to individual examples by filtering via the histogram and scatterplot (see Fig.~\ref{fig:summaryView}) for a case by case exploration, as illustrated in (d)(e). For highly unstable outliers, we often observe the prediction of the original sentences near the decision boundary (e.g., the yellow circle in (f) corresponds to a entailment prediction that is very close to neural), however, some predictions, such as the one illustrated in (d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word ``pile'' is replace by ``heap'' in the hypothesis sentence).
%
}
\label{fig:predictStability}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scenario 4: Explore Relationship Between Grammar and Attention}

There is no simple answer to what exactly is the information attention captures.
Can attention capture grammar structure?

The prediction is Entailment which is wrong. The "green" will be aligned to "green". However they are actually attached to different words in the dependency tree. That is "green" in P is attached to jacket, while "green" in H is attached to tea. They should not be aligned.
And more interestingly is, by forcing the two "green"'s alignment to be zero, the prediction label is corrected to Neutral. This stresses the importance of dependency tree in our tool.

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictStability}
 \vspace{-4mm}
 \caption{
Prediction stability assessment. In (a)(b)(c), we estimate the overall prediction stability (regarding synonymous perturbation) for each type prediction over the entire development set (10k examples). The user can drill down to individual examples by filtering via the histogram or scatterplot (see Fig.~\ref{fig:summaryView}). For highly unstable prediction, we often observe the original prediction is near a decision boundary (e.g., the yellow circle in (e) corresponds to a prediction that is at the boundary between \emph{entailment} and \emph{neutral}), however, some predictions, such as the one illustrated in (d) can alter the prediction quite drastically with minor perturbation (illustrated in $d_1$, where the word ``pile'' is replace by ``heap'' in the hypothesis sentence).
}
\label{fig:predictStability}
\vspace{-2mm}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scenario 4: Explore Relationship Between Grammar and Attention}
The attention computation in the NLI model does not take the grammar structure of the sentences into consideration,
% There is no simple answer to what exactly is the information attention captures.
% Can attention capture grammar structure? 
yet, the attention often can highlight the key structural elements of the sentence. 
Therefore, domain experts want to understand whether attention alone is sufficient to distinguish, 
and more important what kind of benefit can the additional information from grammar parsing help address NLI challenge.

In our tool, we overly the sentence dependency tree with the attention that enables researchers to conduct easy comparable between attention and grammar structure.
As illustrated in Fig.~\ref{fig:depTreeExample}(a), the prediction of the sentence pair (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is \emph{entailment}, which is wrong. We can understand the reasoning by examine the attention, in which we observes that the word \textbf{green} in "\textbf{green} jacket" is aligned to the \textbf{green} in "\textbf{green} tea". Due to such an alignment, the model mistakenly believe the \textbf{green}s are used to describe the same thing (therefore, producing the \emph{entailment} label).  However, as we examine the dependency tree, the two \textbf{green}s are actually attached to different words, i.e., "green" in \textbf{P} is attached to jacket, while "green" in \textbf{H} is attached to tea. Therefore, they should not be aligned in the attention in order to allow the classifier make the right prediction.
%
This experiment demonstrate the potential benefit to include grammar structure in the alignment computation. %This stresses the importance of dependency tree in our tool.
%
And more interestingly is, as illustrated in Fig.~\ref{fig:depTreeExample}(b), by editing the attention and forcing the two \textbf{green}'s alignment to be zero, the prediction label is corrected to \emph{neutral}.

### 2018-03-29 paper/overview.tex
re-arrange figure
update section 6.4

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-3}).
%
As discussed in the task analysis, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the all pair summary view (Section~\ref{sec:allPairs}).

@@

>>>>

@@

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-3}).
%
As discussed in the previous section, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the all pair summary view (Section~\ref{sec:allPairs}).

### 2018-03-29 paper/sentence.tex
update section 6.4

\subsection{Sentence View}
\label{sec:sentence}
% Once an exploration cycle started, the experts need to examine the premise and hypothesis sentence the current example.
As illustrated in Fig.~\ref{fig:sentenceView}, the sentence view shows the premise and hypothesis sentence in (c) and (d).
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces \emph{nouns} and \emph{verbs} by their synonyms in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When the perturbation is applied to either of the sentences, the replaced words are highlighted in blue (see Fig.~\ref{fig:modelPipeline}(e)) to signal the modification made to the original sentence.

@@

At the top left of the view, the two controls \emph{Predict} and \emph{Predict All} (a) correspond to predicting the currently displayed sentence pairs and predict all combinations of perturbed premises and hypotheses, respectively.
%
To avoid the situation where both sentences are perturbed, we ensure only one perturbation per sentence pair (i.e., we use the original premise if the hypothesis is perturbed, or use the original hypothesis if the premise is perturbed).
%%% sentence perturbation %%%%
The previously explored original sentences are stored in the dropdown list (see Fig.~\ref{fig:sentenceView}(b)) that allow the user to revisit examined examples.
Also, the user can type any sentences or modified existing text in the text areas (Fig.~\ref{fig:sentenceView}(c)(d)) to accommodate user-defined input.

>>>>

\subsection{Sentence View}
\label{sec:sentence}
% Once an exploration cycle started, the experts need to examine the premise and hypothesis sentence the current example.
As illustrated in Fig.~\ref{fig:teaser}(B), the sentence view shows the premise and hypothesis sentence.
%
To facilitate the analysis task (\textbf{T1}), we employ an automate sentence perturbation scheme that replaces \emph{nouns} and \emph{verbs} by their synonyms in the wordNet~\cite{Miller1995} (the standard lexical database for NLP applications).
%
When the perturbation is applied to either of the sentences, the replaced words are highlighted in blue to signal the modification made to the original sentence.

@@

At the top left of the view, the two controls \emph{Predict} and \emph{Predict All} correspond to predicting the currently displayed sentence pairs and predict all combinations of perturbed premises and hypotheses, respectively.
%
To avoid the situation where both sentences are perturbed, we ensure only one perturbation per sentence pair (i.e., we use the original premise if the hypothesis is perturbed, or use the original hypothesis if the premise is perturbed).
%%% sentence perturbation %%%%
The previously explored original sentences are stored in the dropdown list (left of the buttons) that allow the user to revisit examined examples.
Also, the user can type any sentences or modified existing text in the sentence display areas to accommodate user-defined input.
